---
title: "QBS 121 Project - BRCAness"
author: "Kenneth MK Mark"
date: "17 May, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Preparation
```{r}
rm(list=ls())
library(ggplot2)
myinf1 = "/Users/kmark/truffula/kmark/BRCAness/iRAS_profile/Curtis_Symbol_GSE54265_BRCA_iRAS_4allsamples.txt"
myinf2A = "/Users/kmark/truffula/kmark/Datasets/Breast_Cancer/Curtis/discovery_Patient_tumor_info.txt"
myinf2B = "/Users/kmark/truffula/kmark/Datasets/Breast_Cancer/Curtis/validation_Patient_tumor_info.txt"
myinf3 <- "/Users/kmark/Google Drive/Dartmouth/2015-2016/Spring_2016/QBS_121/Project/curtis_mut_matrix.txt"
myinf4 <- "/Users/kmark/Google Drive/Dartmouth/2015-2016/Spring_2016/QBS_121/Project/tumor_id_map.txt"

# myinf1 = "/lorax/chenglab/kmark/BRCAness/iRAS_profile/Curtis_Symbol_GSE54265_BRCA_iRAS_4allsamples.txt"
# myinf2A = "/lorax/chenglab/kmark/Datasets/Breast_Cancer/Curtis/discovery_Patient_tumor_info.txt"
# myinf2B = "/lorax/chenglab/kmark/Datasets/Breast_Cancer/Curtis/validation_Patient_tumor_info.txt"

data = read.table(myinf1, sep="\t", header=T, row.names=1)
cnum = ncol(data)/2
data = data[, 1:cnum]
colnames(data) = gsub(".ES", "", colnames(data), fix=T)
cnum = ncol(data)/2
dat1 = data[,1:cnum]
dat2 = data[, (cnum+1):(2*cnum)]
xx = dat1-dat2
colnames(xx) = gsub("_up", "", colnames(dat1))
data = xx*-1 # Switch to HRS Scores

d.inf = read.table(myinf2A, sep="\t", header=T, row.names=1)
v.inf = read.table(myinf2B, sep="\t", header=T, row.names=1)
# sum(d.inf[,"ER.Expr"]=="-" & d.inf[,"PR.Expr"]=="-" & d.inf[,"Her2.Expr"]=="-")
# sum(v.inf[,"ER.Expr"]=="-" & v.inf[,"PR.Expr"]=="-" & v.inf[,"Her2.Expr"]=="-")

info = rbind(d.inf, v.inf)
xx = rep(0, nrow(info))
xx[info[, "last_follow_up_status"]=="d-d.s."] = 1
e.rfs = xx
t.rfs = as.numeric(info[, "T"])
info = cbind(t.rfs, e.rfs, info)
info[, "grade"] = as.numeric(info[, "grade"])
info[, "stage"] = as.numeric(info[, "stage"])
info[, "size"] = as.numeric(info[, "size"])
row.names(info) = gsub("-", ".", row.names(info), fix=T)

## Get the mutational status
sommut <- read.table(myinf3, header = T, quote = "", sep = "\t", row.names = 1)
id.map <- read.table(myinf4, header = T, quote = "", sep = "\t", row.names = 2)

# change the labels to be curtis standard
id.map$sample <- gsub("-", "\\.", id.map$sample)
sommut <- sommut[, which(colnames(sommut) %in% id.map$sample)]
colnames(sommut) <- row.names(id.map)
colnames(sommut) <- gsub("-", "\\.", colnames(sommut))

BRCA1.mut <- sommut[which(row.names(sommut) == "BRCA1"), ]
BRCA1.mut <- as.numeric(BRCA1.mut)
BRCA1.mut[is.na(BRCA1.mut)] <- 0
BRCA1.mut <- ifelse(BRCA1.mut == 0, 0, 1)

BRCA2.mut <- sommut[which(row.names(sommut) == "BRCA2"), ]
BRCA2.mut <- as.numeric(BRCA2.mut)
BRCA2.mut[is.na(BRCA2.mut)] <- 0
BRCA2.mut <- ifelse(BRCA2.mut == 0, 0, 1)
br.stat <- cbind(BRCA1.mut, BRCA2.mut)
row.names(br.stat) <- colnames(sommut)
br.stat <- br.stat[order(row.names(br.stat)), ]

## Add mutational status to info
comm <- intersect(row.names(br.stat), row.names(info))

data <- data[comm, ]
br.stat <- br.stat[comm, ]
info <- info[comm, ]

info <- cbind(info, br.stat)
```

## Is there an association between HRS and Survival?
### Survival analysis - coxph models for all samples together
```{r}
## Normal Coxph models
library(survival)
survreg.pval1 = survreg.pval2 = coxph.pval1 = coxph.pval2 =rep(0, ncol(data))
for(k in 1:ncol(data)){
    mytf <- as.numeric(data[, k])
    xx <- cbind(mytf, info)
    xx <- xx[xx[, "t.rfs"] > 0, ]
    mycox = survreg(Surv(t.rfs, e.rfs)~mytf, xx)
    mycox = summary(mycox)$table
	survreg.pval1[k] = mycox["mytf", "p"]
	mycox = coxph(Surv(t.rfs, e.rfs)~mytf, xx)
	mycox = summary(mycox)
	coxph.pval1[k] = mycox$coefficients[5]
    survreg.qval1 = p.adjust(survreg.pval1, "BH")
    coxph.qval1 = p.adjust(coxph.pval1, "BH")
}

name = colnames(data)
res = data.frame(name, survreg.pval1, survreg.qval1,  coxph.pval1, coxph.qval1)
res = res[order(res[,2]), ]
res

# > res
#             name survreg.pval1 survreg.qval1 coxph.pval1 coxph.qval1
# 4       shBRCA1B  4.725143e-19  2.362572e-18 0.000000000 0.000000000
# 5        shRad51  9.801785e-17  2.450446e-16 0.000000000 0.000000000
# 1         shPTEN  1.042227e-03  1.737045e-03 0.005211619 0.008686032
# 2       shBRCA1A  5.947491e-03  7.434364e-03 0.023889557 0.029861946
# 3 shPTEN.shBRCA1  5.764659e-01  5.764659e-01 0.433668457 0.433668457

## Look at only BRCA1(B) and Rad51

xx <- cbind(RAD51 = as.numeric(data$shRad51), BRCA1 = as.numeric(data$shBRCA1B), info)
xx <- xx[xx[, "t.rfs"]>0, ]
```

### Look at the most significant (RAD51 and BRCA1)
```{r}
## Individual models
fit.rad51 <- coxph(Surv(t.rfs, e.rfs) ~ RAD51, xx)
summary(fit.rad51)

pnorm(summary(fit.rad51)$coefficients[4])

fit.brca1 <- coxph(Surv(t.rfs, e.rfs) ~ BRCA1, xx)
summary(fit.brca1)

pnorm(summary(fit.brca1)$coefficients[4])
```

### Does this remain signficant adjuting for BRCA1 or BRCA2 mutational status
```{r}
## Adjusting for BRCA1 and BRCA2 mutational status
fit.rad51.mut <- coxph(Surv(t.rfs, e.rfs) ~ RAD51 + BRCA1.mut + BRCA2.mut, xx)
summary(fit.rad51.mut)
fit.brca1.mut <- coxph(Surv(t.rfs, e.rfs) ~ BRCA1 + BRCA1.mut + BRCA2.mut, xx)
summary(fit.brca1.mut)
```

### Create survival curves
```{r}
RAD51.mean <- xx$RAD51
RAD51.mean <- ifelse(RAD51.mean > mean(RAD51.mean, na.rm = T), 1, 0)
xx1 <- cbind(RAD51.mean, xx)
curve <- survfit(Surv(t.rfs, e.rfs) ~ RAD51.mean, data = xx1)
tmp <- survdiff(Surv(t.rfs, e.rfs) ~ RAD51.mean, data = xx1)
tmp
plot(curve, mark.time = T, lty=1, lwd=3, col = 2:3, xlab="Time", ylab="P(relapse_free_survival)")
legend(1, 0.2, c("HRS(-) (n=1171)", "HRS(+) (n=735)"), lty=1, lwd=4, col = 2:3)
legend("topright", 1, paste("p.value = 3.3e-16"))
title("Kaplan-Meier Curves Depicting RFS Survival\nfor Curtis Data stratified by HRS (RAD51) Scores")


BRCA1.mean <- xx$BRCA1
BRCA1.mean <- ifelse(BRCA1.mean > 0, 1, 0)
xx1 <- cbind(BRCA1.mean, xx)
curve <- survfit(Surv(t.rfs, e.rfs) ~ BRCA1.mean, data = xx1)
tmp <- survdiff(Surv(t.rfs, e.rfs) ~ BRCA1.mean, data = xx1)
tmp
plot(curve, mark.time = T, lty=1, lwd=3, col = 2:3, xlab="Time", ylab="P(relapse_free_survival)")
legend(1, 0.2, c("HRS(-) (n=1022)", "HRS(+) (n=884)"), lty=1, lwd=4, col = 2:3)
legend("topright", 1, paste("p.value = 5.5e-16"))
title("Kaplan-Meier Curves Depicting RFS Survival\nfor Curtis Data stratified by HRS (BRCA1) Scores")
```

## Breast Cancer subtypes across HRS
### Is there an association between pam50 subtype and HRS?
```{r}
## Is there an association between pam50 subtype and HRS
xx1 <- xx[which(xx$Pam50Subtype != "NC"), ]
fit1 <- lm(RAD51 ~ factor(Pam50Subtype), xx1)

pamr <- xx$Pam50Subtype
pamr <- factor(pamr, levels = c("NC", "Basal", "LumA", "LumB", "Her2", "Normal"))
fit2 <- lm(xx$RAD51 ~ factor(pamr) + BRCA1.mut + BRCA2.mut)
summary(fit2)
```
### Is there an association between ER status and HRS?
```{r}
ER.stat <- xx$ER.Expr
ER.stat <- ifelse(ER.stat == "+", 1, 0)
fit.er <- glm(ER.stat ~ xx$RAD51 + BRCA1.mut + BRCA2.mut, family = "binomial")
summary(fit.er)
```

### an association between PR status and HRS?
```{r}
PR.stat <- xx$PR.Expr
PR.stat <- ifelse(PR.stat == "+", 1, 0)
fit.pr <- glm(PR.stat ~ xx$RAD51 + BRCA1.mut + BRCA2.mut, family = "binomial")
summary(fit.pr)
```

### Association between Her2 and HRS?
```{r}
Her2.stat <- xx$Her2.Expr
Her2.stat <- ifelse(Her2.stat == "+", 1, 0)
fit.her2 <- glm(PR.stat ~ xx$RAD51 + BRCA1.mut + BRCA2.mut, family = "binomial")
summary(fit.her2)
```

### Correlation with somatic mutation count
```{r}
sommut <- data.matrix(sommut)
sommut[is.na(sommut)] <- 0
sommut <- ifelse(sommut==0, 0, 1)
cumsom <- apply(sommut, 2, sum, na.rm = T)
cumsom <- cumsom[order(names(cumsom))]

yy <- data.frame(RAD51 =data$shRad51, cumsom)
yy[, 1] <- ifelse(yy[, 1]>mean(yy[, 1], na.rm = T), "hi", "lo")

# myoutf1 <- "/Users/kmark/Google Drive/Dartmouth/2015-2016/Spring_2016/QBS_121/Project/boxplot.pdf"
# pdf(myoutf1, width = 8, height = 5)
gg1 <- ggplot(yy, aes(factor(RAD51), cumsom))
gg1  + geom_boxplot(notch = T, aes(fill = RAD51)) + ylab("Cumulative Somatic Mutations") + xlab("HR Activity Group")
# dev.off
wilcox.test(yy$cumsom ~ yy$RAD51)
```


## Is this valid in TCGA data?

```{r}
rm(list=ls())
myinf1 = "//Users/kmark/truffula/kmark/BRCAness/iRAS_profile/TCGA_Symbol_GSE54265_BRCA_iRAS.txt"
myinf2 = "~/truffula/kmark/Datasets/Breast_Cancer/TCGA/TCGA_clinical_patient_brca_formatted.rda"

data = read.table(myinf1, sep="\t", header=T, row.names=1)
cnum = ncol(data)/2
data = data[, 1:cnum]
colnames(data) = gsub(".ES", "", colnames(data), fix=T)
cnum = ncol(data)/2
dat1 = data[,1:cnum]
dat2 = data[, (cnum+1):(2*cnum)]
xx = dat1-dat2
colnames(xx) = gsub("_up", "", colnames(dat1))
data = xx*-1
data <- data[order(row.names(data)), ]
load(myinf2)
clin <- clin[order(row.names(clin)), ]

comm <- intersect(row.names(data), row.names(clin))
data <- data[comm, ]
clin <- clin[comm, ]

xx <- cbind(RAD51 = as.numeric(data$shRad51), BRCA1 = as.numeric(data$shBRCA1B), clin)
t.surv <- as.numeric(xx$death_days_to)
e.surv <- xx$vital_status
e.surv <- ifelse(e.surv == "Alive", 0, 1)
xx <- cbind(t.surv, e.surv, xx)
fit.rad51 <- coxph(Surv(t.surv, e.surv) ~ RAD51, data = xx)
summary(fit.rad51)

fit.brca1 <- coxph(Surv(t.surv, e.surv) ~ BRCA1, data = xx)
summary(fit.brca1)
```
